---
title: "Mapping COVID-19 in Missouri"
author: "Christopher Prener, Ph.D."
output: html_document
params:
  date: "Current as of 2020-03-29"
date: '`r params$date`'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(leaflet)
library(sf)
library(here)

snapshot <- st_read(here("data", "county", "daily_snapshot_mo_xl.geojson"), crs = 102003,
                    stringsAsFactors = FALSE) %>%
  st_transform(crs = 4326)
```

## How to Use This Map
This map is fully interactive. Clicking on a county will reveal some details about that place. You can navigate around them just as you would with Google Maps. You can zoom with your mouse or the `plus` and `minus` buttons in the upper-left corner of each map. You may also move the map around by clicking with your mouse and dragging.

## CVOID-19 Confirmed Infection Rates


```{r map, echo = FALSE, out.width = '100%', out.height='800px'}
pal <- colorNumeric("BuGn", snapshot$confirmed_rate)

leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(
    data = snapshot,
    color = "#444444", 
    weight = 1, 
    opacity = 1.0, 
    smoothFactor = 0.5,
    fillOpacity = 0.5,
    fillColor = ~pal(confirmed_rate),
    highlightOptions = highlightOptions(color = "white", weight = 2, bringToFront = TRUE),
    popup = paste("<b>County:</b> ", snapshot$county, "<br>",
                           "<b>Current Confirmed Count:</b> ", snapshot$confirmed, "<br>",
                           "<b>Current Confirmed Rate per 1,000:</b> ", round(snapshot$confirmed_rate, digits = 2))
    )  %>%
    addLegend(pal = pal, values = snapshot$confirmed_rate, opacity = .5, title = "Rate")
```

