---
title: "Mapping COVID-19 in Missouri"
author: "Christopher Prener, Ph.D."
output: html_document
params:
  date: "Current as of 2020-03-29"
date: '`r params$date`'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(leaflet)
library(readr)
library(sf)

source(here::here("source", "functions", "map_breaks.R"))

snapshot <- st_read(here::here("data", "county", "daily_snapshot_mo_xl.geojson"), crs = 102003,
                    stringsAsFactors = FALSE) %>%
  st_transform(crs = 4326) %>%
  mutate(icu_rate = ifelse(hospitals == 0, NA, icu_rate))

zip_snapshot <- st_read(here::here("data", "zip", "daily_snapshot_city_county.geojson"), stringsAsFactors = FALSE)
```

## Introduction
These maps summarize the latest available COVID-19 outbreak data from Johns Hopkins University's [COVID-19 project](https://github.com/CSSEGISandData/COVID-19/tree/master/csse_covid_19_data), the City of St. Louis, and St. Louis County. 

Unlike other interactive maps being used to track the outbreak, the initial three maps include the counties in Illinois and Kansas that are part of the St. Louis and Kansas City metropolitan areas. Kansas City is displayed here as if it were its own county. This is necessary because their public health department is reporting data for the city separate from the four counties that include parts of Kansas City.

The final map is also unique - it includes both the City of St. Louis and St. Louis County on one map and with a shared set of legend values, making direct comparisons possible. It shows Zip Code Tabulation Areas (ZCTAs), which are generalized areas that are roughly equivalent to USPS zip-codes. They are not 100% the same, however, and some homes with a given zip code may fall outside of ZCTA boundaries.

### How to Use These Maps
These maps are fully interactive. Clicking on a county will reveal some details about that place. You can navigate around them just as you would with Google Maps. You can zoom with your mouse or the `plus` and `minus` buttons in the upper-left corner of each map. You may also move the maps around by clicking with your mouse and dragging.

### Caveats
Due to lags in reporting, both at the public health department level and at Johns Hopkins itself, these numbers may lag behind other figures reported in the media. Data are based on confirmed tests and deaths, which are in turn biased by the lack of testing. The extent to which we are increasing testing may also make it seem like there are increases in cases, something known as "surveillance bias." While we are almost certainly seeing spreading infection in Missouri communities, it is important to understand that increased testing is driving some of the observed increases in confirmed cases.

<br>

## Health Care Infrastructure
This first map uses data from the [Kaiser Health Network](https://khn.org/news/as-coronavirus-spreads-widely-millions-of-older-americans-live-in-counties-with-no-icu-beds/) to identify counties (in gray) without any hospitals as well as the number of ICU beds per 1,000 residents in counties that do have hospitals. Keep in mind that some hospitals may have expanded ICU bed capacity in anticipation of increased need. 

For Kansas City, all hospital and ICU bed data have been allocated to Jackson, Clay, Cass, and Platte Counties. If you have a sense of number of beds in Kansas City, MO itself as opposed to the surrounding counties, please feel free to [drop me an email](mailto:chris.prener@slu.edu).

<br>

```{r map-icu, echo = FALSE, out.width = '100%', out.height='600px' , warning=FALSE}
# calculate breaks
bins <- map_bins(snapshot, var = "icu_rate", style = "fisher", classes = 5, dig_lab = 2,
                 round = 2)

# create color palette
pal <- colorBin("PuRd", domain = snapshot$icu_rate, bins = bins)

# map
leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(
    data = snapshot,
    color = "#444444", 
    weight = 1, 
    opacity = 1.0, 
    smoothFactor = 0.5,
    fillOpacity = 0.5,
    fillColor = ~pal(icu_rate),
    highlightOptions = highlightOptions(color = "white", weight = 2, bringToFront = TRUE),
    popup = paste("<b>County:</b> ", snapshot$county, "<br>",
                           "<b>Hospital Count:</b> ", snapshot$hospitals, "<br>",
                           "<b>Baseline ICU Beds:</b> ", snapshot$icu_beds, "<br>",
                           "<b>Baseline ICU Bed Rate per 1,000 Residents:</b> ", round(snapshot$icu_rate, digits = 3))
    )  %>%
    addLegend(pal = pal, values = snapshot$icu_rate, opacity = .5, title = "Rate")
```

<br>

## CVOID-19 Confirmed Infection Rates by County
This map shows confirmed infections as a rate per 1,000 residents. It is important not to map the raw counts themselves, but if you want to see those data, click on a county. You can also view the hospital infrastructure details from the first map for each county by clicking on them.

<br>

```{r map-infections, echo = FALSE, out.width = '100%', out.height='600px'}
# calculate breaks
bins <- map_bins(snapshot, var = "confirmed_rate", style = "fisher", classes = 5, dig_lab = 2,
                 round = 2)
bins[length(bins)] <- 1.66

# create color palette
pal <- colorBin("BuGn", domain = snapshot$confirmed_rate, bins = bins)

# map
leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(
    data = snapshot,
    color = "#444444", 
    weight = 1, 
    opacity = 1.0, 
    smoothFactor = 0.5,
    fillOpacity = 0.75,
    fillColor = ~pal(confirmed_rate),
    highlightOptions = highlightOptions(color = "white", weight = 2, bringToFront = TRUE),
    popup = paste("<b>County:</b> ", snapshot$county, "<br>",
                           "<b>Current Confirmed Count:</b> ", snapshot$confirmed, "<br>",
                           "<b>Current Confirmed Rate per 1,000:</b> ", round(snapshot$confirmed_rate, digits = 2), "<br>",
                           "<b>Hospital Count:</b> ", snapshot$hospitals, "<br>",
                           "<b>Baseline ICU Beds:</b> ", snapshot$icu_beds)
    )  %>%
    addLegend(pal = pal, values = snapshot$confirmed_rate, opacity = .75, title = "Rate")
```

<br>

## CVOID-19 Mortality Rates by County
This map shows confirmed mortality as a rate per 1,000 residents. As with the second map, raw counts of deaths and hospital infrastructure details are available by clicking on individual counties.

<br>

```{r map-mortality, echo = FALSE, out.width = '100%', out.height='600px'}
# calculate breaks
bins <- map_bins(snapshot, var = "mortality_rate", style = "fisher", classes = 5, dig_lab = 2,
                 round = 2)
bins[length(bins)] <- 0.17

# create color palette
pal <- colorBin("YlGn", domain = snapshot$mortality_rate, bins = bins)

# map
leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(
    data = snapshot,
    color = "#444444", 
    weight = 1, 
    opacity = 1.0, 
    smoothFactor = 0.5,
    fillOpacity = 0.75,
    fillColor = ~pal(mortality_rate),
    highlightOptions = highlightOptions(color = "white", weight = 2, bringToFront = TRUE),
    popup = paste("<b>County:</b> ", snapshot$county, "<br>",
                           "<b>Current Mortality Count:</b> ", snapshot$deaths, "<br>",
                           "<b>Current Mortality Rate per 1,000:</b> ", round(snapshot$mortality_rate, digits = 2), "<br>",
                           "<b>Hospital Count:</b> ", snapshot$hospitals, "<br>",
                           "<b>Baseline ICU Beds:</b> ", snapshot$icu_beds)
    )  %>%
    addLegend(pal = pal, values = snapshot$mortality_rate, opacity = .75, title = "Rate")
```

<br>

## CVOID-19 Confirmed Infection Rates by St. Louis ZCTA
This map shows confirmed infections as a rate per 1,000 residents for all ZCTAs with five or more patients. It is important not to map the raw counts themselves, but if you want to see those data, click on a ZCTA. If a ZCTA. straddles the city and the county, and only has confirmed infection numbers in one entity, its estimated partial population will be displayed. Similarly, estimated partial populations for zip codes that straddle outlying counties are used.

<br>

```{r map-infections-zip, echo = FALSE, out.width = '100%', out.height='600px', warning=FALSE}
# calculate breaks
bins <- map_bins(zip_snapshot, var = "confirmed_rate", style = "fisher", classes = 5, dig_lab = 2,
                 round = 2)
# bins[length(bins)] <- 3.62

# create color palette
pal <- colorBin("BuGn", domain = zip_snapshot$confirmed_rate, bins = bins)

# map
leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(
    data = zip_snapshot,
    color = "#444444", 
    weight = 1, 
    opacity = 1.0, 
    smoothFactor = 0.5,
    fillOpacity = 0.75,
    fillColor = ~pal(confirmed_rate),
    highlightOptions = highlightOptions(color = "white", weight = 2, bringToFront = TRUE),
    popup = paste0("<b>Zip:</b> ", zip_snapshot$zip, " <br> ",
                 "<b>Current Confirmed Count:</b> ", zip_snapshot$confirmed, "<br>",
                 "<b>Current Confirmed Rate per 1,000:</b> ", round(zip_snapshot$confirmed_rate, digits = 2), " <br> ",
                 "<b>Population Below Poverty Line:</b> ", round(zip_snapshot$pvty_pct, digits = 2), "% <br> ",
                 "<b>African American Population:</b> ", round(zip_snapshot$blk_pct, digits = 2), "% ")
    )  %>%
    addLegend(pal = pal, values = zip_snapshot$confirmed_rate, opacity = .75, title = "Rate")
```

<br>

## Technical Notes
* The Fisher breaks algorithm (Fisher 1958, Slocum et al. 2005) is used to identify legend categories
* Estimated populations for the the four counties around Kansas City as well as Kansas City itself were created using areal weighted interpolation (see Prener and Revord 2019)
* Estimated populations for partial zip codes are also created using areal weighted interpolation (see Prener and Revord 2019)
