---
title: "Tracking COVID-19 in Missouri"
author: "Christopher Prener, Ph.D."
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    df_print: paged
params:
  date: "Current as of 2020-03-29"
  date_val: "2020-03-29"
  prior_date_val: "2020-03-22"
date: '`r params$date`'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(leaflet)
library(readr)
library(sf)

library(ggplot2)

source(here::here("source", "functions", "map_breaks.R"))
source(here::here("source", "functions", "round_any.R"))

state_data <- read_csv(here::here("data", "state", "state_full.csv")) %>%
  filter(state == "Missouri") %>%
  filter(report_date >= "2020-03-07") %>%
  arrange(desc(report_date))

stl_city_data <- read_csv(here::here("data", "county", "county_full.csv")) %>%
  filter(geoid == "29510") %>%
  filter(report_date >= "2020-03-18") %>%
  arrange(desc(report_date))

stl_county_data <- read_csv(here::here("data", "county", "county_full.csv")) %>%
  filter(geoid == "29189") %>%
  filter(report_date >= "2020-03-07") %>%
  arrange(desc(report_date))

kc_city_data <- read_csv(here::here("data", "county", "county_full.csv")) %>%
  filter(geoid == "29511") %>%
  filter(report_date >= "2020-03-20") %>%
  arrange(desc(report_date))

snapshot <- st_read(here::here("data", "county", "daily_snapshot_mo_xl.geojson"), crs = 4326,
                    stringsAsFactors = FALSE) %>%
  mutate(icu_rate = ifelse(hospitals == 0, NA, icu_rate)) %>%
  mutate(avg_rate = (confirmed_avg/total_pop)*1000)

zip_snapshot <- st_read(here::here("data", "zip", "daily_snapshot_city_county.geojson"), stringsAsFactors = FALSE)
```

## Introduction
These maps summarize the latest available COVID-19 outbreak data from Johns Hopkins University's [COVID-19 project](https://github.com/CSSEGISandData/COVID-19/tree/master/csse_covid_19_data), the New York Time's [COVID-19 project](http://github.com/nytimes/covid-19-data), the City of St. Louis, and St. Louis County. 

For state and county-level data, this site uses the most recent Johns Hopkins University daily release. For all prior days, the New York Times' data are used. The Johns Hopkins data are published on a faster schedule but sometimes are not updated for all states or counties. Conversely, the New York Times' data are published on a slower schedule but have more consistent updates across states and counties. Using the latest Johns Hopkins release allows for faster updates, while using the New York Times' data for all prior days balances our need for recent data with consistency.

### What Makes These Maps Unique?
Unlike other interactive maps being used to track the outbreak, the initial three maps include the counties in Illinois, Kansas, and Oklahoma that are part of Missouri's metropolitan areas. Kansas City itself is also displayed here as if it were its own county. This is necessary because their public health department is reporting data for the city separate from the four counties that include parts of Kansas City.

The final map is also unique - it includes both the City of St. Louis and St. Louis County on one map and with a shared set of legend values, making direct comparisons possible. It shows Zip Code Tabulation Areas (ZCTAs), which are generalized areas that are roughly equivalent to USPS zip-codes. They are not 100% the same, however, and some homes with a given zip code may fall outside of ZCTA boundaries.

### How to Use These Maps
These maps are fully interactive. Clicking on a county will reveal some details about that place. You can navigate around them just as you would with Google Maps. You can zoom with your mouse or the `plus` and `minus` buttons in the upper-left corner of each map. You may also move the maps around by clicking with your mouse and dragging.

### Caveats
Due to lags in reporting, both at the public health department level and at Johns Hopkins itself, these numbers may lag behind other figures reported in the media. 

**As of April 14th, the CDC began requesting data on presumed positive cases.** This means that individuals who physicians believe are infected with COVID-19, but who cannot get access to testing, should be reported. Beginning on April 17th, signs in Missouri's data appeared that suggested at least one county, Moniteau, was reporting both confirmed and presumed positives. This change was confirmed on the Moniteau County's Public Department [COVID-19 website](https://moniteaucovid19.com).

This complicates the data picture, since the data for Moniteau County are now not directly comparable to data for any county in Missouri that is still only reporting those cases that have a confirmed positive test result. **In the short term, this change is going to distort all of the county-level maps and plots on this site.** Please keep that in mind. In an ideal world, all counties would have begun reporting probable cases at the same time in Missouri. Hopefully we are able to get a clearer data picture in the coming days as more counties change their reporting processes.

## Static Maps and Plots {.tabset}
In addition to the interactive maps below, I regularly update a range of static maps and plots that capture the progression of the COVID-19 outbreak in Missouri, several metro areas, and at the ZCTA (zip code) level in St. Louis City and County. The most recent versions of each plot can be viewed by clicking on the links within the tabs below.

### State

| Links |  |  |
|-------------------------|--|--|
| <a href="https://raw.githubusercontent.com/slu-openGIS/covid_daily_viz/master/results/low_res/state/b_confirmed_rate.png" target="_blanl">Confirmed Cases (Rate)</a> | <a href="https://raw.githubusercontent.com/slu-openGIS/covid_daily_viz/master/results/low_res/log_confirmed/a_state.png" target="_blanl">Confirmed Cases (Log)</a> | <a href="https://raw.githubusercontent.com/slu-openGIS/covid_daily_viz/master/results/low_res/state/c_case_fatality_rate.png" target="_blanl">Case Fatality (Rate)</a> |
| <a href="https://raw.githubusercontent.com/slu-openGIS/covid_daily_viz/master/results/low_res/state/d_mortality_rate.png" target="_blanl">Mortality (Rate)</a> |

### County

| Links |  |  |
|-------------------------|--|--|
| <a href="https://raw.githubusercontent.com/slu-openGIS/covid_daily_viz/master/results/low_res/state/a_mo_map.png" target="_blank">Confirmed Cases Map</a> | <a href="https://raw.githubusercontent.com/slu-openGIS/covid_daily_viz/master/results/low_res/log_confirmed/d_missouri.png" target="_blank">Confirmed Cases (Log)</a> | |

### St. Louis Metro

| Links |  |  |
|-------------------------|--|--|
| <a href="https://raw.githubusercontent.com/slu-openGIS/covid_daily_viz/master/results/low_res/stl_metro/a_confirmed_map.png" target="_blank">Confirmed Cases Map</a> | <a href="https://raw.githubusercontent.com/slu-openGIS/covid_daily_viz/master/results/low_res/stl_metro/b_confirmed_plot.png" target="_blank">Confirmed Cases Plot (Rate)</a> | <a href="https://raw.githubusercontent.com/slu-openGIS/covid_daily_viz/master/results/low_res/log_confirmed/b_st_louis.png" target="_blank">Confirmed Cases Plot (Log)</a> | 
| <a href="https://raw.githubusercontent.com/slu-openGIS/covid_daily_viz/master/results/low_res/stl_metro/c_case_fatality_rate.png" target="_blank">Case Fatality (Rate)</a> | <a href="https://raw.githubusercontent.com/slu-openGIS/covid_daily_viz/master/results/low_res/stl_metro/d_case_fatality_plot.png" target="_blank">Mortality (Rate)</a> | |

### Kansas City Metro

| Links |  |  |
|-------------------------|--|--|
| <a href="https://raw.githubusercontent.com/slu-openGIS/covid_daily_viz/master/results/low_res/kc_metro/a_confirmed_map.png" target="_blank">Confirmed Cases Map</a> | <a href="https://raw.githubusercontent.com/slu-openGIS/covid_daily_viz/master/results/low_res/kc_metro/b_confirmed_plot.png" target="_blank">Confirmed Cases Plot (Rate)</a> | <a href="https://raw.githubusercontent.com/slu-openGIS/covid_daily_viz/master/results/low_res/log_confirmed/c_kansas_city.png" target="_blank">Confirmed Cases Plot (Log)</a> | 
| <a href="https://raw.githubusercontent.com/slu-openGIS/covid_daily_viz/master/results/low_res/kc_metro/c_case_fatality_rate.png" target="_blank">Case Fatality (Rate)</a> | <a href="https://raw.githubusercontent.com/slu-openGIS/covid_daily_viz/master/results/low_res/kc_metro/d_case_fatality_plot.png" target="_blank">Mortality (Rate)</a> | |

### St. Louis ZCTA (Zip Code) 

| Links |  |  |
|-------------------------|--|--|
| <a href="https://raw.githubusercontent.com/slu-openGIS/covid_daily_viz/master/results/low_res/zip/a_confirmed_map.png" target="_blank">Confirmed Cases Map</a> | <a href="https://raw.githubusercontent.com/slu-openGIS/covid_daily_viz/master/results/low_res/zip/b_poverty_plot.png" target="_blank">% Poverty Plot</a> | <a href="https://raw.githubusercontent.com/slu-openGIS/covid_daily_viz/master/results/low_res/zip/c_race_plot.png" target="_blank">% African American Plot</a> | 

<br>

## Daily Data {.tabset}
While the [City of St. Louis](https://www.stlouis-mo.gov/covid-19/data/), [St. Louis County](https://stlcogis.maps.arcgis.com/apps/MapSeries/index.html?appid=6ae65dea4d804f2ea4f5d8ba79e96df1), and [Kansas City](https://www.kcmo.gov/city-hall/departments/health/coronavirus-covid-19/covid-19-totals) provide day-by-day tracking of cumulative cases on their respective dashboards, the State of Missouri does not. The following tabs provide daily breakdowns of COVIV data as well as additional statistics not included in the existing dashboards. Please note that the two average columns for reported cases and deaths are both seven-day rolling averages.

### Missouri

```{r missouri-data-table, echo=FALSE, out.width = '100%'}
state_data %>%
  select(report_date, confirmed, new_confirmed, confirmed_avg, deaths, new_deaths, deaths_avg) %>%
  mutate(
    confirmed_avg = round(confirmed_avg, digits = 2),
    deaths_avg = round(deaths_avg, digits = 2)
  ) %>%
  rename(
    `Report Date` = report_date,
    `Cumulative Cases` = confirmed,
    `New Cases` = new_confirmed,
    `Average New Cases` = confirmed_avg,
    `Cumulative Deaths` = deaths,
    `New Deaths` = new_deaths,
    `Average New Deaths` = deaths_avg
  ) -> data_table

data_table
```

### City of St. Louis

```{r stl-city-data-table, echo=FALSE, out.width = '100%'}
stl_city_data %>%
  select(report_date, confirmed, new_confirmed, confirmed_avg, deaths, new_deaths, deaths_avg) %>%
  mutate(
    confirmed_avg = round(confirmed_avg, digits = 2),
    deaths_avg = round(deaths_avg, digits = 2)
  ) %>%
  rename(
    `Report Date` = report_date,
    `Cumulative Cases` = confirmed,
    `New Cases` = new_confirmed,
    `Average New Cases` = confirmed_avg,
    `Cumulative Deaths` = deaths,
    `New Deaths` = new_deaths,
    `Average New Deaths` = deaths_avg
  ) -> data_table

data_table
```

### St. Louis County

```{r stl-county-data-table, echo=FALSE, out.width = '100%'}
stl_county_data %>%
  select(report_date, confirmed, new_confirmed, confirmed_avg, deaths, new_deaths, deaths_avg) %>%
  mutate(
    confirmed_avg = round(confirmed_avg, digits = 2),
    deaths_avg = round(deaths_avg, digits = 2)
  ) %>%
  rename(
    `Report Date` = report_date,
    `Cumulative Cases` = confirmed,
    `New Cases` = new_confirmed,
    `Average New Cases` = confirmed_avg,
    `Cumulative Deaths` = deaths,
    `New Deaths` = new_deaths,
    `Average New Deaths` = deaths_avg
  ) -> data_table

data_table
```

### Kansas City

```{r kansas-city-data-table, echo=FALSE, out.width = '100%'}
kc_city_data %>%
  select(report_date, confirmed, new_confirmed, confirmed_avg, deaths, new_deaths, deaths_avg) %>%
  mutate(
    confirmed_avg = round(confirmed_avg, digits = 2),
    deaths_avg = round(deaths_avg, digits = 2)
  ) %>%
  rename(
    `Report Date` = report_date,
    `Cumulative Cases` = confirmed,
    `New Cases` = new_confirmed,
    `Average New Cases` = confirmed_avg,
    `Cumulative Deaths` = deaths,
    `New Deaths` = new_deaths,
    `Average New Deaths` = deaths_avg
  ) -> data_table

data_table
```

## Health Care Infrastructure
This first map uses data from the [Kaiser Health Network](https://khn.org/news/as-coronavirus-spreads-widely-millions-of-older-americans-live-in-counties-with-no-icu-beds/) to identify counties (in gray) without any hospitals as well as the number of ICU beds per 1,000 residents in counties that do have hospitals. Keep in mind that some hospitals may have expanded ICU bed capacity in anticipation of increased need. 

For Kansas City, all hospital and ICU bed data have been allocated to Jackson, Clay, Cass, and Platte Counties. If you have a sense of number of beds in Kansas City, MO itself as opposed to the surrounding counties, please feel free to [drop me an email](mailto:chris.prener@slu.edu).

<br>

```{r map-icu, echo = FALSE, out.width = '100%', out.height='600px' , warning=FALSE}
# calculate breaks
bins <- map_bins(snapshot, var = "icu_rate", style = "fisher", classes = 5, dig_lab = 2,
                 round = 2)

# create color palette
pal <- colorBin("PuRd", domain = snapshot$icu_rate, bins = bins)

# map
leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(
    data = snapshot,
    color = "#444444", 
    weight = 1, 
    opacity = 1.0, 
    smoothFactor = 0.5,
    fillOpacity = 0.5,
    fillColor = ~pal(icu_rate),
    highlightOptions = highlightOptions(color = "white", weight = 2, bringToFront = TRUE),
    popup = paste("<b>County:</b> ", snapshot$county, "<br>",
                           "<b>Hospital Count:</b> ", snapshot$hospitals, "<br>",
                           "<b>Baseline ICU Beds:</b> ", snapshot$icu_beds, "<br>",
                           "<b>Baseline ICU Bed Rate per 1,000 Residents:</b> ", round(snapshot$icu_rate, digits = 3))
    )  %>%
    addLegend(pal = pal, values = snapshot$icu_rate, opacity = .5, title = "ICU Rate")
```

## Infection Rates by County {.tabset}
This map shows infections as a rate per 1,000 residents. It is important not to map the raw counts themselves, but if you want to see those data, click on a county. You can also view the hospital infrastructure details from the first map for each county by clicking on them or by viewing the data table.

### Map

```{r map-infections, echo = FALSE, out.width = '100%', out.height='600px'}
# calculate breaks
bins <- map_bins(snapshot, var = "confirmed_rate", style = "fisher", classes = 5, dig_lab = 2,
                 round = 2)
bins[length(bins)] <- round_any(x = max(snapshot$confirmed_rate), accuracy = .01, f = ceiling)

# create color palette
pal <- colorBin("BuGn", domain = snapshot$confirmed_rate, bins = bins)

# map
leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(
    data = snapshot,
    color = "#444444", 
    weight = 1, 
    opacity = 1.0, 
    smoothFactor = 0.5,
    fillOpacity = 0.75,
    fillColor = ~pal(confirmed_rate),
    highlightOptions = highlightOptions(color = "white", weight = 2, bringToFront = TRUE),
    popup = paste("<b>County:</b> ", snapshot$county, "<br>",
                           "<b>Current Cumulative Count:</b> ", snapshot$confirmed, "<br>",
                           "<b>Current Cumulative Rate per 1,000:</b> ", round(snapshot$confirmed_rate, digits = 2), "<br>",
                           "<b>Hospital Count:</b> ", snapshot$hospitals, "<br>",
                           "<b>Baseline ICU Beds:</b> ", snapshot$icu_beds)
    )  %>%
    addLegend(pal = pal, values = snapshot$confirmed_rate, opacity = .75, title = "Rate")
```

### Data Table

```{r confirmed-data-table, echo=FALSE, out.width = '100%'}
snapshot %>%
  select(state, county, confirmed, confirmed_rate) %>%
  mutate(
    confirmed_rate = round(confirmed_rate, digits = 2)
  ) %>%
  rename(
    State = state,
    County = county,
    `Cumulative Cases` = confirmed,
    `Rate per 1,000 Residents` = confirmed_rate
  ) %>%
  arrange(State, County) -> data_table

st_geometry(data_table) <- NULL

data_table
```

## Average New Cases by County {.tabset}
This map shows a seven-day rolling average of new cases. For this map, this covers `r params$date_val` back through `r params$prior_date_val`. There is not a threshold for what constitutes a high or low average, but the higher the average number of new cases, the more new spread we can infer. For mapping purposes, these are displayed as a rate per 1,000 residents. As with the prior maps, additional details are available by clicking on each county or on the data table.

### Map

```{r map-confirmed-avg, echo = FALSE, out.width = '100%', out.height='600px', warning=FALSE}
# calculate breaks
bins <- map_bins(snapshot, var = "avg_rate", style = "fisher", classes = 5, dig_lab = 2,
                 round = 2)
bins[length(bins)] <- round_any(x = max(snapshot$avg_rate), accuracy = .01, f = ceiling)

# create color palette
pal <- colorBin("RdPu", domain = snapshot$avg_rate, bins = bins)

# map
leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(
    data = snapshot,
    color = "#444444", 
    weight = 1, 
    opacity = 1.0, 
    smoothFactor = 0.5,
    fillOpacity = 0.75,
    fillColor = ~pal(avg_rate),
    highlightOptions = highlightOptions(color = "white", weight = 2, bringToFront = TRUE),
    popup = paste("<b>County:</b> ", snapshot$county, "<br>",
                           "<b>Current Cumulative Count:</b> ", snapshot$confirmed, "<br>",
                           "<b>7-day Average of New Cases:</b> ", round(snapshot$confirmed_avg, digits = 2), "<br>",
                           "<b>Average New Cases per 1,000 Residents:</b> ", round(snapshot$avg_rate, digits = 2), "<br>",
                           "<b>Hospital Count:</b> ", snapshot$hospitals, "<br>",
                           "<b>Baseline ICU Beds:</b> ", snapshot$icu_beds)
    )  %>%
    addLegend(pal = pal, values = snapshot$avg_rate, opacity = .75, title = "7-day Average Rate")
```


### Data Table

```{r confirmed-avg-data-table, echo=FALSE, out.width = '100%'}
snapshot %>%
  select(state, county, confirmed, confirmed_avg, avg_rate) %>%
  mutate(
    confirmed_avg = round(confirmed_avg, digits = 2)
  ) %>%
  rename(
    State = state,
    County = county,
    `Cumulative Cases` = confirmed,
    `7-day Average New Cases` = confirmed_avg,
    `Average New Cases per 1,000 Residents` = avg_rate
  ) %>%
  arrange(State, County) -> data_table

st_geometry(data_table) <- NULL

data_table
```

## Mortality Rates by County {.tabset}
This map shows mortality as a rate per 1,000 residents. As with the prior maps, raw counts of deaths and hospital infrastructure details are available by clicking on individual counties or on the data table.

### Map

```{r map-mortality, echo = FALSE, out.width = '100%', out.height='600px'}
# calculate breaks
bins <- map_bins(snapshot, var = "mortality_rate", style = "fisher", classes = 5, dig_lab = 2,
                 round = 2)
bins[length(bins)] <- round_any(x = max(snapshot$mortality_rate), accuracy = .01, f = ceiling)

# create color palette
pal <- colorBin("YlGn", domain = snapshot$mortality_rate, bins = bins)

# map
leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(
    data = snapshot,
    color = "#444444", 
    weight = 1, 
    opacity = 1.0, 
    smoothFactor = 0.5,
    fillOpacity = 0.75,
    fillColor = ~pal(mortality_rate),
    highlightOptions = highlightOptions(color = "white", weight = 2, bringToFront = TRUE),
    popup = paste("<b>County:</b> ", snapshot$county, "<br>",
                           "<b>Current Mortality Count:</b> ", snapshot$deaths, "<br>",
                           "<b>Current Mortality Rate per 1,000:</b> ", round(snapshot$mortality_rate, digits = 2), "<br>",
                           "<b>Hospital Count:</b> ", snapshot$hospitals, "<br>",
                           "<b>Baseline ICU Beds:</b> ", snapshot$icu_beds)
    )  %>%
    addLegend(pal = pal, values = snapshot$mortality_rate, opacity = .75, title = "Rate")
```

### Data Table

```{r mortality-data-table, echo=FALSE, out.width = '100%'}
snapshot %>%
  select(state, county, deaths, mortality_rate, case_fatality_rate) %>%
  mutate(
    mortality_rate = round(mortality_rate, digits = 2),
    case_fatality_rate = round(case_fatality_rate, digits = 2)
  ) %>%
  rename(
    State = state,
    County = county,
    `Cumulative Deaths` = deaths,
    `Rate per 1,000 Residents` = mortality_rate,
    `Case Fatality (%)` = case_fatality_rate
  ) %>%
  arrange(State, County) -> data_table

st_geometry(data_table) <- NULL

data_table
```

## Infection Rates by St. Louis ZCTA (Zip Code) {.tabset}
This map shows infections as a rate per 1,000 residents for all ZCTAs with five or more patients. It is important not to map the raw counts themselves, but if you want to see those data, click on a ZCTA or the data table. If a ZCTA straddles the city and the county, and only has reported infection numbers in one entity, its estimated partial population will be displayed. Similarly, estimated partial populations for zip codes that straddle outlying counties are used.

### Map

```{r map-infections-zip, echo = FALSE, out.width = '100%', out.height='600px', warning=FALSE}
# calculate breaks
bins <- map_bins(zip_snapshot, var = "confirmed_rate", style = "fisher", classes = 5, dig_lab = 2,
                 round = 2)
bins[1] <- round_any(x = min(zip_snapshot$confirmed_rate, na.rm = TRUE), accuracy = .01, f = floor)
bins[length(bins)] <- round_any(x = max(zip_snapshot$confirmed_rate, na.rm = TRUE), accuracy = .01, f = ceiling)

# create color palette
pal <- colorBin("BuGn", domain = zip_snapshot$confirmed_rate, bins = bins)

# map
leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(
    data = zip_snapshot,
    color = "#444444", 
    weight = 1, 
    opacity = 1.0, 
    smoothFactor = 0.5,
    fillOpacity = 0.75,
    fillColor = ~pal(confirmed_rate),
    highlightOptions = highlightOptions(color = "white", weight = 2, bringToFront = TRUE),
    popup = paste0("<b>Zip:</b> ", zip_snapshot$zip, " <br> ",
                 "<b>Current Cumulative Count:</b> ", zip_snapshot$confirmed, "<br>",
                 "<b>Current Cumulative Rate per 1,000:</b> ", round(zip_snapshot$confirmed_rate, digits = 2), " <br> ",
                 "<b>Population Below Poverty Line:</b> ", round(zip_snapshot$pvty_pct, digits = 2), "% <br> ",
                 "<b>African American Population:</b> ", round(zip_snapshot$blk_pct, digits = 2), "% ")
    )  %>%
    addLegend(pal = pal, values = zip_snapshot$confirmed_rate, opacity = .75, title = "Rate")
```

### Data Table

```{r zip-data-table, echo=FALSE, out.width = '100%'}
zip_snapshot %>%
  select(zip, confirmed, confirmed_rate, pvty_pct, blk_pct) %>%
  mutate(
    confirmed_rate = round(confirmed_rate, digits = 2),
    pvty_pct = round(pvty_pct, digits = 2),
    blk_pct = round(blk_pct, digits = 2)
  ) %>%
  rename(
    `Zip Code` = zip,
    `Cumulative Cases` = confirmed,
    `Rate per 1,000 Residents` = confirmed_rate,
    `Poverty Rate (%)` = pvty_pct,
    `% African American` = blk_pct
  ) -> data_table

st_geometry(data_table) <- NULL

data_table
```

## Technical Notes
* The Fisher breaks algorithm (Fisher 1958, Slocum et al. 2005) is used to identify legend categories
* Estimated populations for the the four counties around Kansas City as well as Kansas City itself were created using areal weighted interpolation (see Prener and Revord 2019)
* Estimated populations for partial zip codes are also created using areal weighted interpolation (see Prener and Revord 2019)
